Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-08-25 manuscript/manuscript.tex
Modifications in manuscript

In order to make the computation more efficient, we use only second order GLQ for each integration while we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
Given an arbitrary tesseroid, \citet{Uieda2016} redefined its dimensions: $L_\lambda$, $L_\phi$ and $L_r$, where the former ones are the arc-distances measured along its top surface and $L_r$ is the difference between the radii of the top and bottom surfaces.
We must check that the following inequality is satisfied:

@@

\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.

The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
(1) a faster calculation of the distance from the computation point to the tesseroid and 
(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.

In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided and therefore both the accuracy of the algorithm and its computation time.
In this way, the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10 cm), in which case it does not modifies significantly the desired gravitational field.

On the other hand, the accuracy of the method is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.

We have performed similar tests, but now with variable density spherical shell.
From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
The first one constitutes the simplest variation, while the second one represents the fastest varying function. The later one tries to emulate a contact between two different media (e.g. crust and mantle interface).
The analytical solutions for the first two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.

@@

>>>>

In order to make the computation more efficient, we use only second order GLQ for each integration while we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
(1) a faster calculation of the distance from the computation point to the tesseroid and 
(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.

More specifically, \citet{Uieda2016} redefined the dimensions of an arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$. The former ones are the arc-distances measured along its top surface while $L_r$ is the difference between the radii of the top and bottom surfaces.

The adaptive discretization algorithm starts by checking that the following inequality is satisfied:

@@

\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
On the contrary, if the inequality is satisfied, the algorithm computes the desired gravity field using the GLQ approximation.

In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided, and therefore regulates both the accuracy of the algorithm and its computation time.
In this way the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10 cm), in which case it does not modifies significantly the desired gravitational field.

On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy is a matter of study.

In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
We have performed similar tests, but now with variable density spherical shell.
From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
The first one constitutes the simplest variation, while the second one represents the fastest varying function. The later one tries to emulate a contact between two different media (e.g. crust and mantle interface).

@@

\section{Speed Comparison}
